Equivalence Hypergraphs: E-Graphs for Monoidal Theories

### 2023-12-07 sections/combinatorial_semantics.tex
Pushout for EHyp with consistency and proofs

    Then, consider a relation $\consistency^{\mu} \subseteq (V_{\mathcal{G}} + E_{\mathcal{G}}) \times (V_{\mathcal{G}} + E_{\mathcal{G}})$ which we will call a \textit{consistency} relation and which has the following properties.
    \begin{enumerate}
        \label{def:consistency_properties}
        \item For each $v \in V_{\mathcal{G}}$ $v \consistency^{\mu} v$
        \item For each $e \in E_{\mathcal{G}}$ $e \consistency^{\mu} e$
        \item If $v \in s(e)$ then $v \consistency^{\mu} e$
        \item If $v \in t(e)$ then $v \consistency^{\mu} e$
        \item The relation is symmetric and transitive
        \item The relation is defined only for edges $e$ and nodes $v$ such that $[v) = [e) \not = \varnothing$
    \end{enumerate}

%     $\langle (V_\mathcal{G} + E_\mathcal{G}), <_c, \# \rangle$ is an event-structure where $\# \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
%     \begin{enumerate}
%         \label{def:e-hypergraph-req}
%         \item 
%         \[
%     \forall x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G}) \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
%     \] 
%     if $x_1 <_c x_2$ we say that conflict $x_2 \# x_3$ is inherited from the conflict $x_1 \# x_3$. If the conflict is not inherited we say it is \textit{immediate} and denote it with $\#_{\mu}$
%     \item $<_c$ and $\#$ are mutually exclusive
%     \item $\#$ is defined \textit{only} for $x \in (V_\mathcal{G} + E_\mathcal{G})$ such that $[x) \not \eq \varnothing$
%     \item $\#$ should respect connectivity
%     \item \update{Complement of $\conflict_{\mu}$ should be transitive}
%     \item for all distinct $x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G})$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
%     \item for all $x, x' \in (V_\mathcal{G} + E_\mathcal{G})$ $x \#_{\mu} x'$ implies $[x) = [x')$
%         \end{enumerate}

@@

        \item
    if $x_1 \consistency_{\mathcal{F}}^{\mu} x_2$ then $\phi(x_1) \consistency_{\mathcal{G}}^{\mu} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$.
    Or, written functionally, when $x \in E_{\mathcal{F}}$
    \[
        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x)))
        \subseteq
        \consistency_{\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))
    \]
    , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
    And, when $x \in V_{\mathcal{F}}$
    \[
        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x)))
        \subseteq
        \consistency_{\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
    \]
    % Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \conflict_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \conflict_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
\end{enumerate}

@@

    \begin{itemize}
    \item First, we assume that $Z$ is a \textit{discrete} e-hypergraph, i.e. it has no edges.
    \item Second, we assume that $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_i,v_j \in V_{Z}$. \update{This essentially means that $f,g$ do not map nodes to levels of different hierarchy.}
    \label{pushout:assumptions}
    \end{itemize}

@@

    For this we will consider several cases where \update{$f$ and $g$ are mono. I require mono just to simplify things, we can get rid of this requirement if everything else is ok.}.

@@

    \update{We will also build this relation in steps, first we do steps for nodes (1) - (3), then steps (1) - (2) for edges and finally, step (4) for nodes.}

    \begin{remark}
        Essentially, what we build above is that if the relation is undefined, it is inherited from either incident nodes or edges.
    \end{remark}
    \update{If $f$ and $g$ are not mono, the definitions stay the same as $i_{2,E}$ and $i_{1,E}$ are mono.}
    Clearly, all the cases above are disjoint.

@@

    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$.
    We can consider the original consistency relation as a function $\consistency_{X+Y}^{\mu} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}$.
    Quotienting the range of the function gives us $\consistency_{X+Y}'^{\mu} : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote as $\langle []_{V}^{\consistency^{\mu}} []_{E}^{\consistency^{\mu}} \rangle$) applied to the return value of $\consistency_{X+Y}^{\mu}$.
    We then define the new relation for edges as
    \begin{enumerate}
        \item If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
                \begin{align*}
                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E})) =
                    (\langle []_{V}^{\consistency}, []_{E}^{\consistency}\rangle^{*}(\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e))))^{c}
                \end{align*}
        \item $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ in $s([e])$ or $t([e])$ such that $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]) \not = \varnothing$, then
        \[
            \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V}))
        \]
    \end{enumerate}

@@

    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$
    \[
        \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V})) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v'))))^{c}
    \]
    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $e \in E_{X+Y}$ such that either $[v] \in s([e])$ or $[v] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$, then
    \[
        \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}([e]_{E}))
    \]
    \end{enumerate}
        where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$ and $c$ is a reflexive, symmetric and transitive closure.
        % That is, the relation above is the smallest relation that contains $\langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\bigcup_{x' \in E_{X} + E_{Y} | [x']_{E} = [x]_{E}} i_{E_{X} + E_{Y}}(x'))$ and such that if $[x_1] <_{c,\sfrac{X+Y}{\sim}}([x_2])$ and $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_1] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$ then $i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_2] \in \conflict_{\sfrac{X+Y}{\sim}}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}[x_3])$.
        % And similarly for nodes.

@@

    \begin{itemize}
        \item 
        \begin{itemize}

            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_2])$.
              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};i_{1,V}(z)$ and $v_2 = g_{V};i_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
              \begin{align*}
                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_1])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
                %      \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\conflict_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
                %      \text{By homomorphism-ness of $f_{E};i_{1,E}$} \\
                %      = \langle []_{V}^{\conflict}, []_{E}^{\conflict} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\conflict_{Z}(i_{E_{Z}}(z)))
              \end{align*}
              Similarly,
              \[
                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{V_{X} + V_{Y}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
              \]
              and we get well-definedness on the nose.
              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
            \item Now suppose that both $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2)) = \varnothing$ and there exists $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$ such that $[v_1] \in s([e])$ (and automatically $[v_2] \in s([e])$).
                Then,
                  \[
                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
                  \]
            \end{itemize}
        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
            \[
             \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
             \subseteq
             \consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))
             \]
            By definition, 
            \[
                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x)) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x))))^{c}
            \]
            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion.
            If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness is true because empty set is a subset of any set.
            Similarly for edges.
    \end{itemize}

@@

@@

        Once again, $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ mean that this was achieved after performing steps (1) - (2) for edges.

@@

                          $e_1$ and $e_2$ should both have pre-image in either $E_{Y}$ or $E_{X}$. The only possible case then is that $v_2 \in s(e_2)$ and $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$.
                          Then $[v_1') = \varnothing$ and $[v_2) = \varnothing$. Show it is not possible.

@@

                                  Then, necessarily, $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = g_{V};i_{2,V}$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$ and it cant be the case that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}([v_2])}) = [i_{2,E}(e_y)]$.

@@

                          would mean that $e_2$ has a predecessor because $v_2$ does.

@@

                    \item \update{TODO}

@@

@@

@@

                  The last bit also means that $v'' \sim v'$, but because $f$ and $g$ are mono we have at most 2 elements in each equivalence class.

@@

@@

>>>>

    Then, consider a relation $\consistency^{\mu} \subseteq (V_{\mathcal{G}} + E_{\mathcal{G}}) \times (V_{\mathcal{G}} + E_{\mathcal{G}})$ which we will call a \textit{consistency} relation and which has the following properties.
    \begin{enumerate}
        \label{def:consistency_properties}
        \item The relation is an equivalence relation
        \item If $v \in s(e)$ then $v \consistency^{\mu} e$
        \item If $v \in t(e)$ then $v \consistency^{\mu} e$
        \item The relation is defined only for edges $e$ and nodes $v$ such that $[v) = [e) \not = \varnothing$
    \end{enumerate}

%     $\langle (V_\mathcal{G} + E_\mathcal{G}), <_c, \# \rangle$ is an event-structure where $\# \subset (V_\mathcal{G} + E_\mathcal{G}) \times (V_\mathcal{G} + E_\mathcal{G})$ is an irreflexive symmetric relation called \textit{conflict} relation which satisfies the following
%     \begin{enumerate}
%         \label{def:e-hypergraph-req}
%         \item 
%         \[
%     \forall x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G}) \text{, if } x_1 <_c x_2 \text{ and } x_1 \# x_3 \text{ then } x_2 \# x_3
%     \] 
%     if $x_1 <_c x_2$ we say that conflict $x_2 \# x_3$ is inherited from the conflict $x_1 \# x_3$. If the conflict is not inherited we say it is \textit{immediate} and denote it with $\#_{\mu}$
%     \item $<_c$ and $\#$ are mutually exclusive
%     \item $\#$ is defined \textit{only} for $x \in (V_\mathcal{G} + E_\mathcal{G})$ such that $[x) \not \eq \varnothing$
%     \item $\#$ should respect connectivity
%     \item \update{Complement of $\hashtag_{\mu}$ should be transitive}
%     \item for all distinct $x_1, x_2, x_3 \in (V_\mathcal{G} + E_\mathcal{G})$ $x_1 \#_{\mu} x_2$ and $x_2 \#_{\mu} x_3$ implies $x_1 \#_{\mu} x_3$
%     \item for all $x, x' \in (V_\mathcal{G} + E_\mathcal{G})$ $x \#_{\mu} x'$ implies $[x) = [x')$
%         \end{enumerate}

@@

        \item
    if $x_1 \consistency_{\mathcal{F}}^{\mu} x_2$ then $\phi(x_1) \consistency_{\mathcal{G}}^{\mu} \phi(x_2)$, where $x_1, x_2 \in \mathcal{F}$.
    Or, written functionally, when $x \in E_{\mathcal{F}}$
    \[
        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{E_{\mathcal{F}}}(x)))
        \subseteq
        \consistency_{\mathcal{G}}^{\mu}(\phi_{E};i_{E_{\mathcal{G}}}(x))
    \]
    , where $\phi_{V};i_{V_{\mathcal{G}}} : V_{\mathcal{F}} \to V_{\mathcal{G}} + E_{\mathcal{G}}$, and similarly for $\phi_{E};i_{E_\mathcal{G}}$ so that $\langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle : V_{\mathcal{F}} + E_{\mathcal{F}} \to  V_{\mathcal{G}} + E_{\mathcal{G}}$.
    And, when $x \in V_{\mathcal{F}}$
    \[
        \langle \phi_{V};i_{V_{\mathcal{G}}}, \phi_{E};i_{E_{\mathcal{G}}} \rangle^{*}(\consistency_{\mathcal{F}}^{\mu}(i_{V_{\mathcal{F}}}(x)))
        \subseteq
        \consistency_{\mathcal{G}}^{\mu}(\phi_{V};i_{V_{\mathcal{G}}}(x))
    \]
    % Or, on per-component basis, if $i_{E_{\mathcal{F}}}(x_1) \in \hashtag_{\mathcal{F}}(i_{E_{\mathcal{F}}}(x_2))$ then $i_{E_{\mathcal{G}}}(\phi_{E}(x_1)) \in \hashtag_{\mathcal{G}}(i_{E_{\mathcal{G}}}(\phi_{E}(x_2)))$ and similarly for other types of $x_1,x_2$.
\end{enumerate}

@@

    \begin{enumerate}
    \item $Z$ is a \textit{discrete} e-hypergraph, i.e. it has no edges.
    \item $[f_{V}(v_i)) = [f_{V}(v_j))$ and $[g_{V}(v_i)) = [g_{V}(v_j))$ for all $v_i,v_j \in V_{Z}$. \update{This essentially means that $f,g$ do not map nodes to levels of different hierarchy.}
    \item If $[f_{V}(v)) \not = \varnothing$ then $[g_{V}(v)) = \varnothing$ and vice versa.
    \item $\consistency^{\mu}(f(v_i)) = \consistency^{\mu}(f(v_j))$ and $\consistency^{\mu}(f(g_i)) = \consistency^{\mu}(g(v_j))$.
    \item $f$ and $g$ are mono.
    \label{pushout:assumptions}
    \end{enumerate}

@@

    We will consider several cases.

@@

    \update{We will also build this relation in steps, first we do steps for nodes (1) - (3), then step (1) for edges and finally, we will interleave step (2) for edges and step (4) for nodes until fixed point.
            Essentially, this interleaving means that nodes inherit the relation from incident edges and vice versa.
    }
    Clearly, all the cases above are disjoint.

@@

    Let's now define $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}$.
    We can consider the original consistency relation as a function $\consistency_{X+Y}^{\mu} : (V_{X} + V_{Y}) + (E_{X} + E_{Y}) \to 2^{(V_{X} + V_{Y}) + (E_{X} + E_{Y})}$.
    Quotienting the range of the function gives us $\consistency_{X+Y}'^{\mu} : V_{X} + V_{Y} + E_{X} + E_{Y} \to 2^{(\sfrac{V_{X} + V_{Y}}{\sim}) + \sfrac{(E_{X} + E_{Y})}{\sim}}$, which is essentially $\langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}$ (a copairing extended to sequences that we will further denote as $\langle []_{V}^{\consistency^{\mu}} []_{E}^{\consistency^{\mu}} \rangle$) applied to the return value of $\consistency_{X+Y}^{\mu}$.
    We then first define an auxiliary relation $\consistency^{\hashtag}$ for edges as
    \begin{enumerate}
        \item If $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) \not = \varnothing$, then
                \begin{align*}
                    \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) =
                    \langle []_{V}^{\consistency}, []_{E}^{\consistency}\rangle^{*}(\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)))
                \end{align*}
        \item $\consistency_{X+Y}^{\mu}(i_{E_{X} + E_{Y}}(e)) = \varnothing$ and there exists $[v]$ in $s([e])$ or $t([e])$ such that $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]) \not = \varnothing$, then
        \[
            \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e]_{E})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{V_{X} + V_{Y}}([v]_{V}))
        \]
    \end{enumerate}

@@

    \item If there exists $v' \in V_{X+Y}$ such that $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) \not = \varnothing$ and $v' \sim v$
    \[
        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v' \in V_{X} + V_{Y} | [v']_{V} = [v]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')))
    \]
    \item If for all $v' \in V_{X+Y}$ such that $v' \sim v$, $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v')) = \varnothing$ and there exists $e \in E_{X+Y}$ such that either $[v] \in s([e])$ or $[v] \in t([e])$ and $\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) \not = \varnothing$, then
    \[
        \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v]_{V})) = \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{E_{X} + E_{Y}}([e]_{E}))
    \]
    \end{enumerate}
    where $i_*$ are injections into $\sfrac{V_{X} + V_{Y}}{\sim} + \sfrac{E_{X} + E_{Y}}{\sim}$.
    We also build this relation step by step as the relation above: first steps (1) for nodes and edges and then we interleave steps for edges and for nodes.

    We then define
    \[
      \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v])))^{c}
    \]
    \[
      \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])) = (\consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e])))^{c}  
    \]
    where $c$ is a reflexive, symmetric and transitive closure.

@@

    \begin{itemize}
        \item Let us show well-definedness.
        \begin{itemize}

            \item  Assume that $v_1, v_2 \in V_{X+Y}$ and $v_1 \sim v_2$ then $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_1]) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{E_{X} + E_{Y}}[v_2])$.
              $v_1 \sim v_2$ implies there exists $z \in V_{Z}$ such that $v_1 = f_{V};i_{1,V}(z)$ and $v_2 = g_{V};i_{2,V}(z)$ and supposing that for any of $v_i$ $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i)) \not = \varnothing$, then
              \begin{align*}
                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
                %      \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(i_{E_{X} + E_{Y}}(t_1))) = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\hashtag_{X+Y}(f_{E};i_{1,E};i_{E_{X} + E_{Y}}(z)) = \\
                %      \text{By homomorphism-ness of $f_{E};i_{1,E}$} \\
                %      = \langle []_{V}^{\hashtag}, []_{E}^{\hashtag} \rangle^{*}(\langle f_{V};i_{1,V};i_{V_{X} + V_{Y}} ,f_{E};i_{1,E};i_{E_{X} + E_{Y}} \rangle^{*}(\hashtag_{Z}(i_{E_{Z}}(z)))
              \end{align*}
              Similarly,
              \[
                \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = (\langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*}(\bigcup_{v_i \in \{v_1, v_2\}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_i))))^{c}
              \]
              and we get well-definedness on the nose.
              We do not need to check well-definedness for edges as $e_1 \sim e_2$ implies $e_1 = e_2$.
            \item Now suppose that both $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_1)) = \varnothing$ and $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(v_2)) = \varnothing$.
                  This would mean that for all $v_i \in V_{Z}$ $\consistency^{\mu}(f_{V}(v)) = \varnothing$ and $\consistency^{\mu}(g_{V}(z)) = \varnothing$ and automatically
                  \[
                    \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_1])) = \consistency_{\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2]))
                  \]
            \end{itemize}
        \item Putting $\phi = ([]_{V},[]_{E})$ into the definition of a homomorphism gives us
            \[
             \langle []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}, []_{E};i_{\sfrac{E_{X} + E_{Y}}{\sim}} \rangle^{*}(\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
             \subseteq
             \consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))
             \]
            By definition, 
            \[
                \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x)) = \langle []_{V}^{\consistency}, []_{E}^{\consistency} \rangle^{*} (\bigcup_{x' \in V_{X} + V_{Y} | [x']_{V} = [x]_{V}} \consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)))
            \]
            and by recalling that $[]_V^{\consistency} = []_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}$ and similarly for $[]_{E}^{\consistency}$ we get the inclusion $\subseteq \consistency_{\sfrac{X+Y}{\sim}}^{\hashtag}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))$ and the latter is the subset of $\consistency_{\sfrac{X+Y}{\sim}}^{\mu}([]_{V};i_{\sfrac{V_{X} + V_{Y}}{\sim}}(x))$ because it is a closure.
            If $\consistency_{X+Y}^{\mu}(i_{V_{X} + V_{Y}}(x)) = \varnothing$ then homomorphismness is true because empty set is a subset of any set.
            Similarly for edges.
    \end{itemize}

@@

    \label{lemma:choice_of_edges_agnostic}

@@

        We will again show the base case when $[e_1]$ and $[e_2]$ inherit the relation atmost at 1 step.
        That is, $[[e_1]) \not = \varnothing$ and $[[e_2]) \not = \varnothing$ mean that this was achieved after performing steps (1) - (2) for edges.

@@

                          $e_1$ and $e_2$ should both have pre-image in either $E_{Y}$ or $E_{X}$ (because they are from the same connected component).

@@

                                  Then, necessarily, $v'_1 \in s(e_1)$ and $v_1 \sim v'_1$ and $v'_1 = g_{V};i_{2,V}(z)$ and hence $e_1$, $e_2$ have pre-image in $E_{Y}$.
                                  If $v_2 \in s(e_2)$ then it should have a pre-image in $V_{Y}$ and it can not be the case that $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [i_{2,E}(e_y)]$.

@@

                          would mean that $e_2$ has a predecessor because $v_2$ does and $v_2 \in s(e_2)$ because $v_2$ has no pre-image in $V_{Z}$. $[v_2]$ can not inherit the relation because we consider inheritance only upto 1 step: inheritance for $v_2$ would mean that $v$ inherits the relation in more than 1 step.

@@

                          The case when $v_2 = g_{V};i_{2,V}(z_2)$ is symmetric.
                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = [<_{c,X+Y}^{\mu}(i_{V_{X} + V_{Y}})(v_2)]$ is not possible as it would mean that $v_2 \in s(e_2)$, but $[e_2) = \varnothing$.
                    \item $<_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{V_{X} + V_{Y}}{\sim}}([v_2])) = <_{c,\sfrac{X+Y}{\sim}}^{\mu}(i_{\sfrac{E_{X} + E_{Y}}{\sim}}([e']))$ is also not possible as it would mean that $e_2$ inherits the relation in more than 1 step.

@@

                Cases $[e_1) = \varnothing$ and $[e_2) = \varnothing$ and $[e_1) = \varnothing$, $[e_2) \not = \varnothing$ are not possible.

@@

                  This is symmetric to the above.

@@

                  The last bit also means that $v'' \sim v'$, but because $f$ and $g$ are mono we have at most 2 elements in each equivalence class so it reduces to the cases above.

@@

        The proofs for the cases when $[v] \in t([e_1])$ and $[v] \in t([e_2])$ and for the combinations of $t,s$ are similar.

@@

\begin{remark}
    To get the result for arbitrary number of steps we take the above lemmas as base case and procceed by induction on the number of steps.
\end{remark}
